---
title: "Adding Colored Text to ggplot"
date: today
format:
  html:
    code-fold: false
execute:
  warning: false
---

Sometimes using colored text in lieu of a legend can make plots more appealling. This is most often the case when showing data for a small number (2 or 3) different groups.

```{r}
library(tidyverse)
library(ggtext)
library(palmerpenguins) 
library(harrypotter)
library(glue)

theme_set(theme_minimal())

pal <- hp(n = 3, option = "HermioneGranger")
```

For example, imagine we want to plot the bill length vs bill depth of different penguin species. One way to do that would be as follows:

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  labs(
    title = "Bill Length vs Bill Depth by Species"
  ) +
  scale_color_hp(discrete = TRUE, option = "HermioneGranger")
```

This is fine, but we might want to get rid of the legend and explicitly denote species/color mappings in the plot title. We can do that as follows.

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point() +
  labs(
    title = glue("Bill Length vs Bill Depth for <span style='color:{pal[1]}'>Adelie</span>, <span style='color:{pal[2]}'>Chinstrap</span>, and <span style='color:{pal[3]}'>Gentoo</span> Penguins")
  ) +
  scale_color_hp(discrete = TRUE, option = "HermioneGranger") +
  theme(
    legend.position = "none",
    plot.title = element_markdown()
  )
 
```

The key to the above is the `<span>` tags in the title as well as the `plot.title = element_markdown()` argument in the theme. All in all, this requires a little bit more setup, but I think the plot looks  cleaner/more refined.